---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(tibble)
library(dplyr)
library(ggplot2)
library(dslabs)
library(scales)


library(QuantPsyc)
library(MASS)
library(data.table)
library(tidyr)
library(GGally)


```


1
glimpse(gapminder)

1.a

```{r}
gapminder%>%
  arrange(country)%>%
  select(country, year,gdp)%>%
  mutate(previous_year_gdp= lag(gdp, order_by = country))
```



1.b



```{r}
gapminder%>%
  select(continent, year,gdp)%>%
  group_by(continent, year) %>%
  summarise_at(vars(gdp),sum,na.rm=TRUE)%>%
  mutate(previous_year_gdp= lag(gdp))%>%
  mutate(change=(gdp/previous_year_gdp-1))
```




```{r}

glimpse(gapminder)

gapminder%>%
  select(continent, country,year,gdp)%>%
  group_by(country)%>%
  mutate(previous_year_gdp= lag(gdp),change=(gdp/previous_year_gdp-1))%>%
  group_by(continent, year) %>%
  summarise(avg_growth=mean(change,na.rm=TRUE))%>%
  ggplot(aes(x=year,y=avg_growth,group=continent,color=continent))+
  geom_line()+
  scale_y_continuous(name="Annual GDP change", limits=c(-0.07,0.1))
```
1.1.2

```{r}
gapminder %>%
  arrange(country)%>%
  select(country, year,gdp,region)%>%
  mutate(py_gdp = lag(gdp))%>%
  mutate(growth=gdp/py_gdp-1)%>%
  group_by(region)%>%
  mutate(mgr=mean(growth,na.rm=TRUE))%>%
  mutate(diff = growth-mgr,na.rm=TRUE)%>%
  group_by(country)%>%
  summarise(sum_diff=sum(diff,na.rm=TRUE))%>%
  arrange(sum_diff)
```
1.1.3


```{r}
gapminder %>%
  select(country, year,gdp,region)%>%
  group_by(country)%>%
  mutate(py_gdp = lag(gdp))%>%
  mutate(growth=gdp/py_gdp-1,na.rm=TRUE)%>%
  group_by(region)%>%
  mutate(mgr=mean(growth,na.rm=TRUE))%>%
  mutate(bool_var = growth>mgr,na.rm=TRUE)%>%
  group_by(country)%>%
  summarise(summa=sum(bool_var,na.rm = TRUE))%>%
  arrange(desc(summa))



```
1.2.1

1.2.1.1.a&b

```{r}
gapminder%>%
  select(country,year,infant_mortality)%>%
  filter (country=="China")%>%
  ggplot( aes(x=infant_mortality))+
  geom_histogram(binwidth = 3)
              
```
1.2.1.2.a
```{r}

  ggplot(gapminder, aes(x=life_expectancy))+
  geom_histogram(binwidth = 3)
              
```
1.2.1.2.b
```{r}
gapminder %>%
  select(country, year,life_expectancy)%>%
  group_by(country,year)%>%
  mutate(bool_var = life_expectancy<40,na.rm=TRUE)%>%
  group_by(country)%>%
  summarise(summa=sum(bool_var,na.rm = TRUE))%>%
  arrange(desc(summa))%>%
  mutate(n1=sum(summa,na.rm=TRUE))
```
```{r}
gapminder %>%
  select(country, year,life_expectancy)%>%
  group_by(country,year)%>%
  mutate(bool_var = life_expectancy>80,na.rm=TRUE)%>%
  group_by(country)%>%
  summarise(summa=sum(bool_var,na.rm = TRUE))%>%
  arrange(desc(summa))%>%
  mutate(n1=sum(summa))
```

1.2.1.3

You probably want to use coord_cartesian instead of lims. Unexpected things can happen when you're fiddling around with the limits on histograms, because a fair bit of fiddly transformations have to happen to get from your raw data to the actual histogram.

So now you're wondering, how the heck did that happen, right?

Well, when you tell ggplot that you want 10 bins and the x limits go from 86 to 195, the histogram algorithm tries to create ten bins that span that actual range. That's why it's trying to create bins down below 100 even though there's no data there.

And then further oddities can happen because the bars may extend past the nominal data range (the xmin and xmax values), since the bar widths will generally encompass a little above and a little below your actual data at the high and low ends.

coord_cartesian will adjust the x limits after all this processing has happened, so it bypasses all these little quirks.
```{r}

  ggplot(gapminder, aes(x=life_expectancy,na.rm=FALSE))+
  geom_histogram(binwidth = 1)+
    coord_cartesian(
      xlim = c(10,80),
      ylim = c(0,400),
      expand = TRUE,
      default = FALSE,
      clip = "on"
)

              
```
1.2.2.1

```{r}


  ggplot(gapminder, aes(x=life_expectancy,na.rm=FALSE))+
  geom_bar()+
    coord_cartesian(
      xlim = c(40,40.5),
      ylim = c(0,3),
      expand = TRUE,
      default = FALSE,
      clip = "on"
)
```
1.2.2.2
if you sum over a column with na result is na
if you take a mean over  a column with ne the result is na

you can fix by using (,na.rm=TRUE)


2.1

glimpse(diamonds)


```{r}

glimpse(diamonds)
summary(diamonds)
ggpairs(diamonds)

```

1.a It is the carat (the weight), the heavier the diamond, the more valuable it is.  There is a strong correlation, 0,922. (x,y,and z) are also strongly correlated to carat as the are the dimensions of the diamond and the bigger they are the heavier it is)

1.b
To show the relationship between price and cut we can look at the quantiles:
```{r}
diamonds %>%
  group_by(cut) %>%
  summarise(n=n(), 
            mean= mean(price), 
            median=median(price), 
            Q1= quantile(price,0.25),
            Q3= quantile(price,0.75))
```
It should be noted that the weight distribution is very different for different cuts.  Therefore the 

```{r}


    
cor(as.integer(diamonds$cut),diamonds$price)

diamonds%>%
  group_by(cut)%>%
  mutate(icut=as.integer(cut))


glimpse(diamonds)
```
2.1.b

```{r}
cor(as.integer(diamonds$cut),diamonds$price)
```
2.1.c. fair = 1, Good = 2, Very good = 3, Premium = 4 and Ideal = 5
2.1.c shows the there is very l.ow negative correlation between the cut and the price of diamonds.  This is because the weight/carat is far more important than the cut

```{r}

cor(diamonds$carat,diamonds$price)

```
This is best seen by noting that the correlation between carats and price is very high or posititve 0.92.


2.2.a

```{r}
color_cut<-
  count(diamonds, color,cut)
  
color_cut

```
2.2.b

  
```{r}
cut_color<-
  count(diamonds, cut, color)%>%
  group_by(cut,color)
  
cut_color
```
  
2.2.c






```{r}
ggplot(data=diamonds,aes(x=color, group=cut))+
  geom_bar(aes(y=..prop..),stat="count")+
  facet_wrap(~cut)

```

2.3 (þetta átti ekki að gera svona)


```{r}
ggplot(data = diamonds, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.3)))
```


```{r}
ggplot(data = diamonds, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_number(carat, 20)))
```

2.3.a
```{r}
ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(colour = cut_width(carat, 1)), binwidth = 500)
```

```{r}
ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(colour = cut_number(carat, 5)), binwidth = 500)
```



2.4

```{r}
ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price), alpha = 1 / 100)
```
2.5

2.6

  
```{r}
ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price), alpha = 1 / 100)+
  facet_wrap(~cut)
```
  

2.7

```{r}
ggplot(data = diamonds) +
geom_point(mapping = aes(x = x, y = y)) +
coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))


```
This is a very good way to see the outliers as the points may appear to normal when looked at seperately as x or y, but the x,y combination shows that they are outliers.


```{r}
ggplot(data = diamonds) +
geom_bin2d(mapping = aes(x = x, y = y)) +
coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))

```
Although the above picture shows that the range of the bins on the y and x axes is substantial it does not really show us where the outliers are. For example at 9-10 carats there are hardly any outliers and the relathinship is almost x=y (lenght = width) this is not easy to see from the binned plot
